---
title: "Creating a speed heatmap (AKA brainscan) using PORTAL Freeway Data"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Freeway Speed Heatmap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The purpose of this vignette is to demonstrate how to use the `freewaySpeedMap` function to create a "brainscan" plot for a corridor within PORTAL's available freeway data. 

##Initialization
First, we load the neccesary packages to create the database connection and review our options for freeway corridors. 
```{r,warning=FALSE,message=FALSE}
#Package imports
require(portalr)
require(rjson)
require(DBI)
require(RPostgreSQL)

db_cred = fromJSON(file="db_credentials.json")
con <- dbConnect(dbDriver("PostgreSQL"), 
                 host=db_cred$db_credentials$db_host, 
                 port= 5432, 
                 user=db_cred$db_credentials$db_user, 
                 password = db_cred$db_credentials$db_pass,
                 dbname=db_cred$db_credentials$db_name)
#Fetch corridor metadata
corridors  = dbGetQuery(con,"SELECT * FROM public.corridors")
#Sort corridors by ID
corridors = corridors[order(corridors$corridorid),]
#Subset to new corridor definitions (i.e. corridor ID > 7000 & < 8000)
corridors[corridors$corridorid >7000 & corridors$corridorid <8000,
          c("corridorid","shortname","longname")]
```

## Plotting 
Let's try a couple of different corridors during different time periods.

This will plot for the 'Portland Central City to Vancouver - NB' corridor during August for weekdays. 
```{r,fig.width=7,fig.height=4}
freewaySpeedMap(con= con,startDate = "2015-08-01", endDate = "2015-08-31",
                weekdays = TRUE)
```

