---
title: "Querying, Aggregating, and Exploring PORTAL Freeway Data"
author: "Bryan Blanc"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Freeway Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The purpose of this vignette is to provide examples of how to use `portalr` to query, aggregate and explore freeway data from PORTAL, Portland's regional transportation data archive. 

## Querying

Analysts can used the `freewayData()` function to request data for one or more detectors over a range of time.

Before you can query the database, you must create a database connection object, as demonstrated below. This step requires the use of a file called 'db_credentials.json', which contains your database authentication information. 
```{r,warning=FALSE,message=FALSE}
#Package imports
require(portalr)
require(rjson)
require(DBI)
require(RPostgreSQL)
require(plyr)

db_cred = fromJSON(file="db_credentials.json")
con <- dbConnect(dbDriver("PostgreSQL"), 
                 host=db_cred$db_credentials$db_host, 
                 port= 5432, 
                 user=db_cred$db_credentials$db_user, 
                 password = db_cred$db_credentials$db_pass,
                 dbname=db_cred$db_credentials$db_name)
#Initialize some metadata
detectors = dbGetQuery(con,"SELECT * FROM public.detectors")
stations= dbGetQuery(con,"SELECT * FROM public.stations")
```
After creating the database connection object, you can begin to start querying data. There are three parameters used to create a basic PORTAL query using `portalr`:

- Detector IDs (`dets`): A vector of detector IDs for which you would like to query data. These can be found within the PORTAL station metadata. 
- Start Time (`startTime`): A string indicating what date/time to begin your query in the YYYY-MM-DD or YYYY-MM-DD HH:MM:SS formats. 
- End Time (`endTime`): A string indicating what date/time to end your query in the YYYY-MM-DD or YYYY-MM-DD HH:MM:SS formats. 

An example query is demonstrated below, and `head()` is used to display the first few rows of the table.

```{r}
#I-205 SB @ Clackamas Hwy
stas = c(1100) 
dets = detectors$detectorid[detectors$stationid %in% stas]
data = freewayData(con = con, dets = dets , startTime = '2015-07-06',endTime = '2015-07-10')
head(data)
```

Typically, this very fine resolution of data (20-second) is not that useful to transportation analysts. That's why the next step we're going to talk about is aggregation. 

## Aggregating
Analysts can use the `aggTime()` function to aggregate data to some resolution of time more useful than 20-second. A few examples are given below.

I want to separate by lane number in my aggregation, so I can join lanenumber in from the ‘detectors’ table in the ‘public’ schema of PORTAL.
```{r}
joined = join(data,detectors, by="detectorid")
agg = aggTime(unAgg=joined,aggVars = "lanenumber",timeCut = "15 min",acrossDays= FALSE)
head(agg)
```
In the below aggregation, I am going to cut the data by hour, and than aggregate it by hour of day across all the days in the dataset.
```{r}
agg_hod = aggTime(unAgg=joined,aggVars = "lanenumber",timeCut = "1 hour",acrossDays = TRUE)
head(agg_hod)
```
This data can be easily visualized using the ggplot2 package. In the first plot, I can see the oscillation in mean hourly traffic speed by lane. 

## Visualizing
```{r, fig.width=7, fig.height=3}
require(ggplot2)
#Lane number needs to be converted from integer to factor for plotting
agg$lanenumber = factor(agg$lanenumber)
ggplot(agg,aes(x=period,y=speed,group =lanenumber, colour = lanenumber))+
  geom_line()+scale_colour_discrete(name="Lane Number")+
  scale_x_datetime()+
  ggtitle("I-205 SB @ Clackamas Hwy \n(Hourly speeds over 2015-07-06 to 2015-07-10)")+
  ylab("Mean Speed (mph)")+xlab("")
```

I can visualize the hour of day graph using ggplot and setting up a few of the labels.

```{r, fig.width=7, fig.height=3}
require(ggplot2)
require(scales)
#Lane number needs to be converted from integer to factor for plotting
agg_hod$lanenumber = factor(agg_hod$lanenumber)

ggplot(agg_hod,aes(x=time,y=speed,group =lanenumber, colour = lanenumber))+
  geom_line()+scale_colour_discrete(name="Lane Number")+
  ggtitle("I-205 SB @ Clackamas Hwy \n(Mean hourly speed over 2015-07-06 to 2015-07-10)")+
  ylab("Mean Speed (mph)")+xlab("")+
  scale_x_datetime(labels = date_format("%I:%M %p", tz=Sys.timezone()))
```

## Last Steps

Always close all your connections when you are done.

```{r}
dbDisconnect(con)
```

