---
title: "Analyzing Specific Issues"
author: "Bryan Blanc"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Specific Issues}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The purpose of this vignette is to demonstrate how the functions in this package were used to diagnose several PORTAL data quality issues. 

## Initialization
First, we must set up our workspace. This includes attaching the neccesary packages (including `portalr`!), setting up a database connection, and loading the station and detector metadata tables. 
```{r,message=FALSE, warning=FALSE}
#Load neccesary libraries
require(portalr)
require(RPostgreSQL)
require(DBI)
require(plyr)
require(ggplot2)
require(lubridate)
require(rjson)

#Connect to Portal db, make sure you VPN into PSU CECS network
db_cred = fromJSON(file="db_credentials.json")
con <- dbConnect(dbDriver("PostgreSQL"), 
                 host=db_cred$db_credentials$db_host, 
                 port= 5432, 
                 user=db_cred$db_credentials$db_user, 
                 password = db_cred$db_credentials$db_pass, 
                 dbname=db_cred$db_credentials$db_name)

#Read in relational Tables
stations= dbGetQuery(con,"SELECT * FROM public.stations")
detectors = dbGetQuery(con,"SELECT * FROM public.detectors")
```

## Issue 20: I-84 WB west of Grand – Very High Lane Volumes
The radar detector located at I-84 Westbound west of Grand Ave. was noted by an ODOT staff member to have very high lane volumes. This prompted the specific investigation of the detector at this station. 

### Querying the data
First, we have to figure out which detectors belong to this station. A 'station' is a geographic location where information is being gathered by one or more 'detectors'. In this case, a Wavetronix radar sensor is gathering data about three lanes of traffic, which are seperated into individual 'detectors'. The station ID was grabbed from the station metadata for this station, and the detector IDs were then pulled from the detector metadata based on the station ID. The data was then queried based on the detector IDs and a specified time range; the month of July was used for the example here. 
```{r}
sta = 3126
dets = detectors$detectorid[detectors$stationid %in% sta]
startDate = "2015-07-01"
endDate = "2015-07-31"
raw = freewayData(con,dets,startDate,endDate)
```

### Aggregating the data
```{r}
joined = join(raw,detectors,by="detectorid")
joined$dow = weekdays(joined$starttime,abbreviate = TRUE)
mf = subset(joined,joined$dow != "Sat" & joined$dow != "Sun")
agg = aggTime(mf,"lanenumber","hour",hod=FALSE)
head(agg)
```

### Plotting the data
```{r,fig.width=7,fig.height=4, message=FALSE, warning=FALSE}
agg$hod=hour(as.POSIXct(agg$period))
timeLabs = c("3 AM","6 AM","9 AM","12 PM","3 PM","6 PM","9 PM")
timeTicks = seq(3,21,3)
ggplot(agg,aes(x=hod,y=volume,group=lanenumber,colour=lanenumber))+
  geom_point(alpha=0.3)+geom_smooth()+xlab("Hour of Day")+
  ylab("Volume (vph)")+
  scale_x_continuous(breaks = timeTicks,labels = timeLabs)+
  scale_y_continuous(limits=c(0,3000))+
  ggtitle("Weekday volumes plotted by hour of day and lane \n for I-84 WB W of Grand")+
  scale_colour_discrete(name="Lane Number")
```



